% 3. Proposed Algorithms: Multi-weighted Contraction Hierarchies
%  - Proposed solution consists of two algorithms for two 
%    separate stages: graph preprocessing and the search query  

\section{Proposed Algorithms: Multi-weighted Contraction Hierarchies}
\label{secAlgorithms}

To address the problem of finding the Pareto set of optimal paths in multi-weighted directed multi-graphs, we propose Multi-weighted Contraction Hierarchies (MCH). 
MCH approach consists of two separate stages, each solved by a separate algorithm: contracted graph construction (MCHp) and query (MCHq). 
Both algorithms are generalisation of Contraction Hierachies for single-weighted undirected graphs \cite{geisberger2008contraction}.

\begin{itemize}
  \item {\em MCHp: Preprocessing algorithm} takes a multi-weighted directed multigraph as the input and returns an extended graph with the same set of nodes, but with additional edges. It contracts all the nodes one by one, adding new edges called \emph{shortcuts}, while preserving all the Pareto optimal shortest path lengths. 
Nodes will have a rank ($rank(v) \in \{0, \dots n-1\}$) corresponding to their order of contraction.
The resulting preprocessed graph has all the original nodes $V$, edges $E$ and new set of shortcuts $E_+$.
Each shortcut $(u,v)_i$ corresponds to a path in the original graph $p_{u,v}$, while $w((u,v)_i) = w(p_{u,v})$ and it keeps a note which is later used to reconstruct the original path.
  \item {\em MCHq: Query algorithm} uses the preprocessed graph to find the Pareto set of all
the shortest paths between two nodes $s$ and $t$. MCHq runs a search algorithm in both directions -- from $s$ to $t$ and from $t$ to $s$. After that, it joins both partial paths into the path in the preprocessed graph and finally reconstructs the path in the original graph. 
\end{itemize}

\input{algorithmsStage1.tex}
\input{algorithmsStage2.tex}


\subsection{Correctness and Completeness}
\label{secFormalResults}

Let us now prove that the query algorithm returns the whole Pareto set of shortest paths.

\begin{lemma}\label{lemmaCorrectness}
%[Lemma]
Set of the edges found by the query algorithm is the Pareto set of shortest paths.
\end{lemma}
\begin{proof}[Proof of Lemma \ref{lemmaCorrectness}]
\ \\ 
\begin{itemize}
\item[$\supseteq$] 
    Let $p_{s,t}$ be an arbitrary path from the Pareto set $P(s,t)$. We show that $p_{s,t}$ is found by the query algorithm. 

Suppose that $t$ is the node with the highest rank in the path $p_{s,t}$. 
Then path $p_{s,t}$ should be found by forward search
and together with path $p_{t,t}$, $l(p_{t,t})=0$ 
found by backward search it represents an original path.
If $p_{s,t}$ is a single edge, then the forward search will find it.
If $p$ is a node with the second highest rank, then we denote $v$ the
node with the third highest rank. 
By the time when $w$ was contracted, all the other nodes had been contracted
and $s,t$ was adjacent to $v$. So the shortcut $(s,t)_j$ was created, and forward search used it.
If $v \ne s$ is the node with the second highest rank on the path,
then the sub-path $p_{s,v}$ is found by the forward search by induction and
then sub-path $p_{v,t}$ is also found by the forward search.

The case when $rank(s)$ is the highest rank on path $p_{s,t}$ is quite similar. 

Let $v \ne s \ne t$ be a node with the highest rank on the path $p_{s,t}$. 
Sub-path $p_{s,v} \in P(s,v)$.\\ 
If it wasn't, than there would be a path $p'_{s,v} \in P(s,v)$ which dominates $p_{s,v}$. 
Then $p'_{s,v}$ could substitute $p_{s,v}$ in $p_{s,t}$. The resulting path $p'_{s,t}$ 
would dominate $p_{s,t}$ which would contradict $p_{s,t} \in P(s,t)$.\\
%\todo[inline]{predchadzajuca veta mi nedava uplne zmysel. opravit formulaciu prosim.}
Similarly, $p_{v,t} \in P(v,t) $. Path $p_{s,v}$ is found by the forward search end path
$p_{v,t}$ by the backward search.
\item[$\subseteq$] 
Each path found by multi-weighted CH corresponds to the path in the original graph. 
The query algorithm filters paths which are dominates by other. So the every path
found by the query algorithm is one of the non-dominated shortest path in original graph.
\end{itemize}
\end{proof}

