% 1. Introduction:
% 
%  (background) - Common problem: Multi-objective shortest path problems (MOSP)
%  - Briefly explain the problem in natural language (no formalisms) - How this
%  graph problem relates to real world (path planning) - The easier variant of
%  the problem, SPP, can be solved using Contraction Hierarchies (CH) - In 1-2
%  natural language sentences, explain the contraction hierarchies approach
%  
%  (problem) - CH has good results (reference), but it was no used for MOSP yet
%  
%  (solution) - in this paper, we show how CH can be used for MOSP - [describe
%  the solution] 
%  
%  (outline) - paper outline. contents of all the sections

\section{Introduction} \label{secIntroduction}

Multi-objective shortest path problem (MOSP) is a fundamental problem from
graph theory with many real world applications. Algorithms solving the problem
need to find the Pareto set of shortest paths between two nodes in a
multi-weighted directed or undirected graphs or multi-graphs. 

A typical example of real-world application of MOSP is automated planning of
paths within a transport network represented as a graph, while optimizing
criteria such as travel time, distance, price, comfort or compatibility with
traveller's preferences. This problem arises from both travel industry and
daily life activities. Trip planning applications, such as car navigation
systems need efficient algorithms for for multi-criteria problems in very large
transport 
% networks~\cite{veneti2016time}. tahle citace je podle me uplne mimo
networks~\cite{duque2015exact}


%  (problem) - CH has good results (reference), but it was no used for MOSP yet
%  CSH

One of the common methods used to speed up the shortest path search  is called
Contraction Hierarchies (CH)~\cite{geisberger2008contraction}. Contraction
Hierarchies can be used to find a shortest path more efficiently than previous
routing approaches and is used in many advanced routing
techniques~\cite{delling2009engineering}. 

The problem is that Contraction Hiearchies approach is not applicable to multi-weighted
graphs and therefore cannot be used for MOSP, even though the method was
recently used for a similar problems. For example, Funke and Storand used CH for Constrained Shortest Path Problem (CSPP)~\cite{pugliese2013survey,funke2013polynomial}. Similarly to Multi-objective shortest path, Constrained Shortest Path algorithms also search
for shortest paths in multi-weighted graphs, but instead of complete Pareto
sets, they only return a single path with the best linear combination of
weights. Before that, Artmeier et al. applied it when searching for shortest paths for electric vehicles, where they optimized for
time, while also considering battery state~\cite{artmeier2010shortest}. 


%  (solution) - in this paper, we show how CH can be used for MOSP - [describe
%  the solution] 

In this paper, we present an algorithm that generalizes the idea behind Contraction
Hierarchies approach to efficiently solve MOSP. Our algorithm, called
Multi-weighted Contraction Hierarchies (MCH), is able to find the whole Pareto
set of shortest paths between a pair of nodes in a multi-weighted directed
multi-graph. In addition to CH, we also draw inspiration from Multicriterial
Label-Setting (MLS) algorithm~\cite{martins1984multicriteria}. 

Even though some methods that are more recent and more advanced than CH exist (for example Hub Labelling by Abraham et al.~\cite{abraham2011hub}), we decided to extend CH because it can be relatively easily generalized for multi-objective search. 


%For vehicle or bicycle routing could be used namoa \todo{citace} which is
%optimal for direct (no preparation) questions.  \todo[inline]{to je state of
%the art a u te si nejsem jista jestli jsem si naimplementovala dobre a tak ji
%nemam ve vysledcich, nevim jestli ji davat tady.} -> radsej nie. 

%There could be improvemenents when we allows preprocessing algorithms. 

Higway dimension  \cite{abraham2016highway} is parametr which is low on road
network whit only on criteria and it could be used for upper bound of time
complexity of search algorithms. We could generalise their definition of 
sparse hitting set. We use vector of single-weighted diameter for each objective
as a multi-weighted diameter. Their analyze of CH query and preparation time
could by applied for multi-weighted case.
%, generalize highway dimension in orther to found complexity bound also for our problem.  
%Intuice v pozadí: Important vertices plati nejen čase - dokazano v ale věřím
%že bude funguvat i v namořské výšce a pohodlnosti cest Tj delší důležitější
%cesty mají hezké vlastnosti


%Fully dynamic speed-up techniques for multi-criteria shortest path searches in
%time-dependent networks.\cite{berger2010fully}

% bidirectional heuristic (spis nepouzitelna) \cite{holte2016bidirectional}
% future work hub labeling \cite{abraham2012hierarchical}
% Prece jen se da pouzit. Staci prubezne scitat a udrzovat bag s nejkrasnima cestama
% algoritmus zastavime pokud vsechny nove otevirane labely tvori cesty dominovate nema nalezenyma

%  (outline) - paper outline. contents of all the sections
After the introduction of required terminology in Section
\ref{secPreliminaries}, we describe our two-stage solution in Section
\ref{secAlgorithms} -- MHC consists of two algorithms: one for 
graph preprocessing and one for queries. In Section
\ref{secExperiments}, we describe our experiments on DIMACS New York City road
network benchmarks and compare the performance to an alternative approach.
Finally, in Section \ref{secFormalResults}, we formally prove our algorithm's
correctness.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \todo[inline]{Tato sekcia potrebuje podstatne viac referencii.}
%Este k tym referenciam v uvode - mam pocit, ze by slo aspon pridat niekolko prikladov ako kto pouzil CH na SPP a na aky real-world problem to bolo aplikovane. Pripadne ake ine techniky sa pouzivaju pri 
